Local Spectral Clustering of Density Upper Level Sets

### 2019-11-10 arxiv/local_spectral_density_clustering_arxiv.tex
Updated per Siva comments

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. The ability of spectral algorithms to perform clustering -- community detection-- in this context is well-understood. \sbcomment{Add reference to McSherry -- spectral partitioning...} In particular, \citet{rohe2011} upper bound the fraction of nodes misclassified by a spectral algorithm for the high-dimensional (large number of blocks) SBM,  \citet{lei2015} extend these results to the sparse (low average degree) regime, and \citet{balakrishnan2011} analyze the misclassification rate when the block model exhibits some hierarchical structure
\sbcomment{add reference to Cris Moore, Newman et. al., and Liza Levina, Bickel, Sarkar have a recent paper that looks at this same hierarchical model}. The framework we consider, where nodes correspond to data points sampled from an underlying density, and edges between nodes are formed based on geometric proximity, is quite different than the SBM, and therefore so is our analysis.

@@

A substantial portion of our technical effort involves estimating graph-theoretic connectivity properties of empirical density clusters in the neighborhood graph $G_{n,r}$. In particular, we prove an upper bound on the mixing time of a random walk run only over the subset of nodes in $G_{n,r}$ which fall within a density cluster $\Cset$. To do so, we rely on a series of seminal works upper bounding the mixing time of \emph{geometric random walks} (see \citet{vempala2005} for a comprehensive review.)  This study was initiated by \citet{dyer1991}, who used geometric random walks as a fundamental subroutine to efficiently compute volumes of high-dimensional convex bodies. These results are improved in \citet{lovasz1990,kannan97,kannan06}, who show, inter alia, that the bounds on mixing time can be sharpened by avoiding so-called ``start-penalties''. As we discuss further in the sequel, these improvements are crucial to our work. Following the work of \citet{abbasi-yadkori2016,abbasi-yadkori2016a}, we extend these results to hold for Lipschitz deformations of convex sets. Additionally, we relate the mixing time of these (continuous-space) geometric random walks to the mixing time of random walks over (discrete) neighborhood graphs. \sbcomment{Refer to and add a couple of lines about the paper of Zhu et al.}

Finally, it is worth mentioning that density clustering and level set estimation are well-studied problems. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} \sbcomment{Add reference to Kpotufe, von Luxburg} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present our first main result: a bound on the volume of the symmetric set difference between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[\Xbf]$. 
In this theorem, and hereafter, we let 
$c, c_i > 0$ represent universal constants, and 
$c_i(\Pbb,r)$ represent constants which depend only on $\Pbb$ and $r$ but not on the sample size $n$; where possible, we keep track of all constants in our proofs.
\begin{theorem}
  \label{thm:volume_ssd}
  Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$. If Algorithm \ref{alg: ppr} is \emph{well-initialized} with respect to $\Cset$, then 
  %there exists a universal constant $c > 0$, constants in sample size $c_1(\Pbb,r)$ - $c_6(\Pbb,r)$, and 
  there exists a set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ of large volume 
  with $\vol(\Csig[\Xbf]^g; G_{n,r}) \geq \frac{1}{2}\vol(\Csig[\Xbf];G_{n,r})$ such that the following statement holds: let Algorithm~\ref{alg: ppr} be run with any seed node $v \in \Csig[\Xbf]^g$. Then, for any $n$ large enough so that
  \begin{equation*}
  n \geq \max\left\{(\log n)^{\max\{\frac{3}{d},1\}}\left(\frac{1}{c_2(\Pbb,r)}\right)^d, c_3(\Pbb,r)\right\},
  \end{equation*}
  the volume of the symmetric set difference is upper bounded
  \begin{equation}
    \label{eqn:volume_ssd}
    \Delta(\Csig[\Xbf], \Cest) \leq c \cdot \kappa(\Cset) \cdot \vol_{n,r}(\Csig[\Xbf]),
  \end{equation}
  with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-(c_1(\Pbb,r) + c_6(\Pbb,r))n}$. 
\end{theorem}
\sbcomment{At this point not sure why you need so many different constants. I would in general assume that $d$ is a constant. Can you not just write for 
$n \geq c_1(\Pbb,r)  (\log n)^{\max\{\frac{3}{d},1\}}$, with probability at least $1 - c_2(\Pbb,r)/n$ blah...?}
The proof of Theorem \ref{thm:volume_ssd} is deferred to Appendix~\ref{sec:proof_of_volume_ssd}. 
This result shows that the volume of the symmetric set difference \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.
\sbcomment{Not completely clear what you are doing above Zhu et. al. Perhaps emphasize again that the technical work is in showing that $\tau_u$ and $\phi_u$ are valid upper bounds on the normalized cut and mixing time, and once
this is established the result follows from Zhu et al.? (unless I am missing something) }

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$, and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized and if 
  %, there exists a universal constant $c >  0$ and constants in sample size $c_1(\Pbb,r)$-$c_6(\Pbb,r)$ such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{(\lambda_{\sigma} r^d
      \nu_d)^2}{\vol_{\Pbb,r}(\Csig)}, 
  \end{equation}
  then $\Cest$ satisfies~\eqref{eqn: consistent_density_cluster_recovery} with respect to $\Cset$ with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-(c_1(\Pbb,r) + c_6(\Pbb,r))n}$. Therefore, $\Cest$ is a consistent estimator of $\Cset$ in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}
\sbcomment{Same suggestion about removing unnecessary constants in the statement of the theorem. You can leave them in the proof.}
\noindent Some remarks are in order: 
 \begin{enumerate}  
 \item We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in an upper bound on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound in~\eqref{eqn: ppr_gap}. 
 \item While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring that the graph $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is common in spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{enumerate}

\subsection{Approximate PPR vector} 
In practice, exactly solving the system of equations~\eqref{eqn:
  ppr_vector} to compute the PPR vector 
  may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\varepsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \varepsilon \Dbf_{uu}\leq
\pbf^{(\varepsilon)}(u) \leq \pbf(u).  
\end{equation}
For a sufficiently small choice of $\varepsilon$, the 
application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:volume_ssd} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\varepsilon)}$}. We
formally state and prove this in Corollary~\ref{cor: appr} in 
Appendix~\ref{sec:appr}.
\sbcomment{Maybe we should discuss approximate PPR briefly in the Background section -- and just state the corollary here (not in a separate section).}

@@

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda > 0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies
  Assumptions \ref{asmp: bounded_density}--\ref{asmp: low_noise_density}, 
  \ref{asmp: bounded_volume} for some $\sigma > 0$. Then there exists a universal constant $c_1 > 0$ and a constant in sample size $c_1(\Pbb,r)$ such that
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}}
  \end{equation}
  with probability at least $1 - 3\exp\{-nc_1(\Pbb,r)\}$. 
\end{theorem}
\noindent {\bf Remark: } 
Observe that the diameter $\rho$ is absent from Theorem \ref{thm:
    conductance_upper_bound}, in contrast to the condition number
  $\kappa(\Cset)$, which worsens (increases) as $\rho$ increases. This
  reflects established wisdom regarding spectral partitioning
  algorithms more generally \citep{guattery1995, hein2010}, albeit newly applied
  to the density clustering setting. It suggests that if the diameter $\rho$ is
  large, PPR may fail to recover $\Csig[\Xbf]$ even when $\Cset$ is
  sufficiently well-conditioned to ensure $\Csig[\Xbf]$ has a small normalized
  cut in $G_{n,r}$. This intuition will be supported by simulations in Section
  \ref{sec: experiments}.

@@

\begin{theorem}
  \label{thm: mixing_time_upper_bound}
  Fix $\lambda > 0$, and assume that \smash{$\Cset \in \Cbb_f(\lambda)$}
  satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}
  for some $\sigma > 0$. Additionally assume that the radius $0 < r < \sigma/2\sqrt{d}$. Then there exist universal constants $c_2$ and $c_3$, and constants in sample size $c_2(\Pbb,r)$ - $c_6(\Pbb,r)$ such that the following statement holds: for any
  \begin{equation*}
  n \geq \max\left\{(\log n)^{\max\{\frac{3}{d},1\}}\left(\frac{1}{c_2(\Pbb,r)}\right)^d, c_3(\Pbb,r)\right\}
  \end{equation*}
  the mixing time $\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$ satisfies
  \begin{equation} 
    \label{eqn: mixing_time_upper_bound}
    \tau_{\infty}(G_{n,r}[\Csig[\Xbf]]) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{ \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
  \end{equation}
  with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-c_6(\Pbb,r)n}$.
\end{theorem}
\sbcomment{Same comment about constants.}
The proof of Theorem \ref{thm: mixing_time_upper_bound} relies heavily on
analogous mixing time bounds developed for the mixing time of a
continuous-space ``ball walk'' over convex sets. To the best of our knowledge, our
result is the first bound on the mixing time of random walks
over neighborhood graphs that is independent of $n$, the number of vertices.

@@

To show a lower bound for density clustering using PPR, we exhibit a hard case: that is, a distribution $\Pbb$ for which PPR is unlikely to recover a density cluster. Let $\mathcal{C}_{\sigma}^{(0)}$, $\mathcal{C}_{\sigma}^{(1)}$, and $\mathcal{C}_{\sigma}^{(2)}$ be rectangles in $\Reals^2$, 
\begin{equation*}
\mathcal{C}_{\sigma}^{(0)} = \biggl[-\frac{\sigma}{2}, \frac{\sigma}{2}\biggr] \times \biggl[-\frac{\rho}{2}, \frac{\rho}{2}\biggr], \quad \mathcal{C}_{\sigma}^{(1)} = \mathcal{C}_{\sigma}^{(0)} - \set{(\sigma,0)}, \quad \mathcal{C}_{\sigma}^{(2)} = \mathcal{C}_{\sigma}^{(0)} + \set{(\sigma,0)} \tag{$0 < \sigma < \rho$}
\end{equation*}
and let $\mathbb{P}$ be the mixture distribution over $\mathcal{X} = \mathcal{C}_{\sigma}^{(0)} \cup \mathcal{C}_{\sigma}^{(1)} \cup \mathcal{C}_{\sigma}^{(2)}$ given by
\begin{equation*}
\mathbb{P} = \frac{1 - \epsilon}{2} \Psi_1 + \frac{1 - \epsilon}{2} \Psi_2 + \frac{\epsilon}{2} \Psi_0,
\end{equation*}
where $\Psi_m$ is the uniform distribution over $\Csig^{(m)}$ for $m = 0,1,2$. 
The density function $f$ of $\Pbb$ is simply
\begin{equation}
\label{eqn:lb_density}
f(x) = \frac{1}{\rho\sigma}\left(\frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(1)}) + \frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(2)}) + \frac{\epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(0)})  \right)
\end{equation}
so that for any $\epsilon < \lambda < (1 - \epsilon)/2$, $\mathbb{C}_{f}(\lambda) = \set{\mathcal{C}_{\sigma}^{(1)}, \mathcal{C}_{\sigma}^{(2)}}$. 
\sbcomment{Maybe a figure would be helpful.}

@@

\begin{remark}Technically, the rectangles $\Csig^{(1)}$ and $\Csig^{(2)}$ are not $\sigma$-expansions due to their sharp corners. To fix this, one can 
%either modify the upper bound (specifically, Lemma~\ref{lem: expansion_volume} in the proof of Theorem~\ref{thm: conductance_upper_bound}) to hold with respect to rectangles of width $\sigma$, or 
change the sets $\Csig^{(1)}$ and $\Csig^{(2)}$ to be rectangles with rounded corners. With some additional bookkeeping, using rounded rectangles does not change our lower bound significantly so we ignore this technicality in our 
subsequent discussion.
\end{remark}

@@

\begin{remark}
	It is worth pointing out that the above conclusions are reliant on specific (albeit reasonable) ranges and choices of input parameters, which in some instances differ between the upper and lower bounds. We suspect that our lower bound continues to hold even when choosing input parameters as dictated by our upper bound; however, proving this is currently beyond our technical grasp, and is a matter for future work. 
\end{remark}

>>>>

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. The ability of spectral algorithms to perform clustering -- community detection-- in this context is well-understood, dating back to \citet{mcsherry2001} who gives conditions under which the entire community structure can be recovered. In more recent work, \citet{rohe2011} upper bound the fraction of nodes misclassified by a spectral algorithm for the high-dimensional (large number of blocks) SBM and \citet{lei2015} extend these results to the sparse (low average degree) regime. Relatedly, \citet{clauset08,balakrishnan2011,li2018}, analyze the misclassification rate when the block model exhibits some hierarchical structure. The framework we consider, where nodes correspond to data points sampled from an underlying density, and edges between nodes are formed based on geometric proximity, is quite different than the SBM, and therefore so is our analysis.

@@

Our analysis naturally builds on a few of the aforementioned theoretical analyses of PPR. For an arbitrary graph $G$ and subset $S \subset G$, \citet{andersen2006} relate the quality of the PPR cluster $\hat{C}$ to the normalized cut functional $\Phi(S;G)$. While this analysis is tight, it fails to account for improvements possible when the cluster $S$ is additionally assumed to be internally well-connected, which is intuitively a more favorable case for clustering. Building on this intuition, \citet{zhu2013} assume that the subgraph $G[S]$ is internally well-connected-- measured by a functional such as mixing time of a random walk over $G[S]$-- and prove upper bounds on $\vol(\widehat{C} \vartriangle S;G)$. Both of these analyzes hold with respect to a approximate form of PPR (APPR) which can be efficiently computed.

We apply these results to our setting by estimating these functionals in the particular case of $G = G_{n,r}$ and $S = \mathcal{C}[X]$. In particular, we prove an upper bound on the mixing time of a random walk run only over the subset of nodes in $G_{n,r}$ which fall within a density cluster $\Cset$. To do so, we rely on a series of seminal works upper bounding the mixing time of \emph{geometric random walks} (see \citet{vempala2005} for a comprehensive review.)  This study was initiated by \citet{dyer1991}, who used geometric random walks as a fundamental subroutine to efficiently compute volumes of high-dimensional convex bodies. These results are improved in \citet{lovasz1990,kannan97,kannan06}, who show, inter alia, that the bounds on mixing time can be sharpened by avoiding so-called ``start-penalties''. As we discuss further in the sequel, these improvements are crucial to our work. Following the work of \citet{abbasi-yadkori2016,abbasi-yadkori2016a}, we extend these results to hold for Lipschitz deformations of convex sets. Additionally, we relate the mixing time of these (continuous-space) geometric random walks to the mixing time of random walks over (discrete) neighborhood graphs.

Finally, it is worth mentioning that density clustering and level set estimation are well-studied problems. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013,kpotufe11} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present our first main result: a bound on the volume of the symmetric set difference between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[\Xbf]$. 
In this theorem, and hereafter, we let 
$c, c_i > 0$ represent universal constants, and 
$c(\Pbb,r), c_i(\Pbb,r) > 0$ represent constants which may depend on $\Pbb,\lambda$ and $r$ but not on the sample size $n$; where possible, we keep track of all constants in our proofs.
\begin{theorem}
  \label{thm:volume_ssd}
  Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$. If Algorithm \ref{alg: ppr} is \emph{well-initialized} with respect to $\Cset$, then 
  %there exists a universal constant $c > 0$, constants in sample size $c_1(\Pbb,r)$ - $c_6(\Pbb,r)$, and 
  there exists a set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ of large volume 
  with $\vol(\Csig[\Xbf]^g; G_{n,r}) \geq \frac{1}{2}\vol(\Csig[\Xbf];G_{n,r})$ such that the following statement holds: let Algorithm~\ref{alg: ppr} be run with any seed node $v \in \Csig[\Xbf]^g$. Then, for any $n$ large enough so that
  \begin{equation*}
  n \geq c_1(\Pbb,r)\cdot (\log n)^{\max\{\frac{3}{d},1\}}
  \end{equation*}
  the volume of the symmetric set difference is upper bounded
  \begin{equation}
    \label{eqn:volume_ssd}
    \Delta(\Csig[\Xbf], \Cest) \leq c \cdot \kappa(\Cset) \cdot \vol_{n,r}(\Csig[\Xbf]),
  \end{equation}
  with probability at least $1 - \frac{c_2(\Pbb,r)}{n}$. 
\end{theorem}
The proof of Theorem \ref{thm:volume_ssd} is deferred to Appendix~\ref{sec:proof_of_volume_ssd}. We reiterate that the primary technical work involved in proving Theorem~\ref{thm:volume_ssd} involves showing that $\tau_{u}$ and $\phi_u$ are valid upper bounds on the normalized cut and mixing time; once this is established, the result follows more or less straightforwardly from~\citet{zhu2013}.
This result establishes that the volume of the symmetric set difference \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$, and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized with respect to $\Cset$ and if 
  %, there exists a universal constant $c >  0$ and constants in sample size $c_1(\Pbb,r)$-$c_6(\Pbb,r)$ such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{(\lambda_{\sigma} r^d
      \nu_d)^2}{\vol_{\Pbb,r}(\Csig)}, 
  \end{equation}
  then the following statement holds: there exists a set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ of large volume 
  with $\vol(\Csig[\Xbf]^g; G_{n,r}) \geq \frac{1}{2}\vol(\Csig[\Xbf];G_{n,r})$ such that if Algorithm~\ref{alg: ppr} is run with any seed node $v \in \Csig[\Xbf]^g$,  then for any $n$ large enough so that
  \begin{equation*}
  n \geq c_1(\Pbb,r)\cdot (\log n)^{\max\{\frac{3}{d},1\}}
  \end{equation*}$\Cest$ satisfies~\eqref{eqn: consistent_density_cluster_recovery} with respect to $\Cset$ with probability at least $1 - \frac{c_2(\Pbb,r)}{n}$. Therefore, $\Cest$ is a consistent estimator of $\Cset$ in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}
\noindent Some remarks are in order: 
 \begin{enumerate}  
 \item We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in an upper bound on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound in~\eqref{eqn: ppr_gap}. 
 \item While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring that the graph $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is common in spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{enumerate}

\subsection{Approximate PPR vector} 
In practice, exactly solving the system of equations~\eqref{eqn:
  ppr_vector} to compute the PPR vector 
  may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\varepsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \varepsilon \Dbf_{uu}\leq
\pbf^{(\varepsilon)}(u) \leq \pbf(u).  
\end{equation}
For a sufficiently small choice of $\varepsilon$, the 
application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:volume_ssd} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\varepsilon)}$}.

\begin{corollary}
	\label{cor: appr}
	Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ be a $\kappa$-well-conditioned cluster for some $\sigma > 0$. Choose input parameters $r, \alpha, v, (L,U)$ to be well-initialized in the sense of \eqref{eqn: initialization}, additionally set 
	\begin{equation}
	\label{eqn: appr_parameter}
	\varepsilon = \frac{1}{25}\vol_{\Pbb,r}(\Csig)
	\end{equation}
	and modify Algorithm \ref{alg: ppr} to compute the a\pprspace vector $\pbf^{(\varepsilon)}_v$ rather than the exact \pprspace vector $\pbf_v$, with resulting output $\Cest$. Then
	\begin{enumerate}
		\item The upper bound on symmetric set difference metric $\Delta(\Cest,\Csig[\Xbf])$ given by \eqref{eqn:volume_ssd} is still valid.
		\item If \ref{asmp: C'_bounded_density} and \eqref{eqn: kappa_ub} hold, 
		then $\Cest \subseteq \Xbf$ is a consistent estimator for $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}.
	\end{enumerate}
\end{corollary}

@@

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda > 0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies
  Assumptions \ref{asmp: bounded_density}--\ref{asmp: low_noise_density}, 
  \ref{asmp: bounded_volume} for some $\sigma > 0$. Then
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}}
  \end{equation}
  with probability at least $1 - 3\exp\{-nc_3(\Pbb,r)\}$. 
\end{theorem}
\noindent {\bf Remark: } 
Observe that the diameter $\rho$ is absent from Theorem \ref{thm:
    conductance_upper_bound}, in contrast to the condition number
  $\kappa(\Cset)$, which worsens (increases) as $\rho$ increases. This
  reflects established wisdom regarding spectral partitioning
  algorithms more generally \citep{guattery1995, hein2010}, albeit newly applied
  to the density clustering setting. It suggests that if the diameter $\rho$ is
  large, PPR may fail to recover $\Csig[\Xbf]$ even when $\Cset$ is
  sufficiently well-conditioned to ensure $\Csig[\Xbf]$ has a small normalized
  cut in $G_{n,r}$. This intuition will be supported by simulations in Section
  \ref{sec: experiments}.

@@

\begin{theorem}
  \label{thm: mixing_time_upper_bound}
  Fix $\lambda > 0$, and assume that \smash{$\Cset \in \Cbb_f(\lambda)$}
  satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}
  for some $\sigma > 0$. Additionally assume that the radius $0 < r < \sigma/2\sqrt{d}$. Then the following statement holds: for any
  \begin{equation*}
  n \geq c_1(\Pbb,r)\cdot(\log n)^{\max\{\frac{3}{d},1\}}
  \end{equation*}
  the mixing time $\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$ satisfies
  \begin{equation} 
    \label{eqn: mixing_time_upper_bound}
    \tau_{\infty}(G_{n,r}[\Csig[\Xbf]]) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{ \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
  \end{equation}
  with probability at least $1 - \frac{c_4(\Pbb,r)}{n}$.
\end{theorem}
The proof of Theorem \ref{thm: mixing_time_upper_bound} relies heavily on
analogous mixing time bounds developed for the mixing time of a
continuous-space ``ball walk'' over convex sets. To the best of our knowledge, our
result is the first bound on the mixing time of random walks
over neighborhood graphs that is independent of $n$, the number of vertices.

@@

To show a lower bound for density clustering using PPR, we exhibit a hard case: that is, a distribution $\Pbb$ for which PPR is unlikely to recover a density cluster. Let $\mathcal{C}_{\sigma}^{(0)}$, $\mathcal{C}_{\sigma}^{(1)}$, and $\mathcal{C}_{\sigma}^{(2)}$ be rectangles in $\Reals^2$, 
\begin{equation*}
\mathcal{C}_{\sigma}^{(0)} = \biggl[-\frac{\sigma}{2}, \frac{\sigma}{2}\biggr] \times \biggl[-\frac{\rho}{2}, \frac{\rho}{2}\biggr], \quad \mathcal{C}_{\sigma}^{(1)} = \mathcal{C}_{\sigma}^{(0)} - \set{(\sigma,0)}, \quad \mathcal{C}_{\sigma}^{(2)} = \mathcal{C}_{\sigma}^{(0)} + \set{(\sigma,0)} \tag{$0 < \sigma < \rho$}
\end{equation*}
and let $\mathbb{P}$ be the mixture distribution over $\mathcal{X} = \mathcal{C}_{\sigma}^{(0)} \cup \mathcal{C}_{\sigma}^{(1)} \cup \mathcal{C}_{\sigma}^{(2)}$ given by
\begin{equation*}
\mathbb{P} = \frac{1 - \epsilon}{2} \Psi_1 + \frac{1 - \epsilon}{2} \Psi_2 + \frac{\epsilon}{2} \Psi_0,
\end{equation*}
where $\Psi_m$ is the uniform distribution over $\Csig^{(m)}$ for $m = 0,1,2$. 
The density function $f$ of $\Pbb$ is simply
\begin{equation}
\label{eqn:lb_density}
f(x) = \frac{1}{\rho\sigma}\left(\frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(1)}) + \frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(2)}) + \frac{\epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(0)})  \right)
\end{equation}
so that for any $\epsilon < \lambda < (1 - \epsilon)/2$, $\mathbb{C}_{f}(\lambda) = \set{\mathcal{C}_{\sigma}^{(1)}, \mathcal{C}_{\sigma}^{(2)}}$. Figure~\ref{fig:fig0} shows $f$ for two different choices of $\epsilon, \sigma$ and $\rho$.

\begin{figure}[tb]
	\centering

	\includegraphics[width=0.4\textwidth]{lower_bound_plots/lower_bound_1}
	\includegraphics[width=0.4\textwidth]{lower_bound_plots/lower_bound_2}

	\caption{\it\small The density $f$ given by~\eqref{eqn:lb_density}, for two different choices of $\epsilon$ and $\sigma$. Left: $\epsilon = .3,\sigma = .1$; Right: $\epsilon = .2,\sigma = .2$. In both $\rho = 1$. } 
	\label{fig:fig0}

\end{figure}

@@

\textbf{Remark:}~~Technically, the rectangles $\Csig^{(1)}$ and $\Csig^{(2)}$ are not $\sigma$-expansions due to their sharp corners. To fix this, one can 
%either modify the upper bound (specifically, Lemma~\ref{lem: expansion_volume} in the proof of Theorem~\ref{thm: conductance_upper_bound}) to hold with respect to rectangles of width $\sigma$, or 
change the sets $\Csig^{(1)}$ and $\Csig^{(2)}$ to be rectangles with rounded corners. With some additional bookkeeping, using rounded rectangles does not change our lower bound significantly so we ignore this technicality in our 
subsequent discussion.

@@

\textbf{Remark:}~~It is worth pointing out that the above conclusions are reliant on specific (albeit reasonable) ranges and choices of input parameters, which in some instances differ between the upper and lower bounds. We suspect that our lower bound continues to hold even when choosing input parameters as dictated by our upper bound; however, proving this is currently beyond our technical grasp, and is a matter for future work.
